Time Series Modeling of Solute Transport in an H\textsubscript{II} Phase Lyotropic Liquid Crystal Membrane

### 2019-09-07 Ben_Manuscripts/stochastic_transport/Outline.tex
MRS3 comments.

  % MRS1: I would start with something more central to the content of the paper - how do solutes move in complex separations membranes?  Let people know right from the start what the paper is about.  This paragraph should still be in the intro, but start with something that captures the main question of the paper. 
  % BJC2: This paragraph will need more tweaking but is the gist of what I'm thinking
  % BJC2: I'm not sure if this is exactly what you meant though
%  Making sense of solute motion in complex separations membranes is central to 
%  understanding transport mechanisms and formulating membrane design principles.
  Mathematical descriptions of transport in complex separations membranes are a powerful 
  way to understand mechanisms and formulate membrane design principles.
  \begin{itemize}
    \item The complexity of a model generally parallels the complexity of the 
    transport mechanism at hand, as well as the desired outputs.
    \item In dense homogeneous membranes, the solution-diffusion model can extract
    diffusion and partition coefficients and has successfully predicted solute transport
    rates.
    \item Analogously, pore-flow models yield predictions of diffusion coefficients
    and solute transport rates in nanoporous membranes.~\cite{wijmans_solution-diffusion_1995}
    \item Single particle tracking allows us to further characterize more complex
    diffusive behavior.
    \item Finally, Molecular simulation provides a framework for studying both single
    particle dynamics and bulk transport properties with atomic resolution.
    \item All of these approaches facilitate speculation on the molecular origins of
    separations and suggest experiments that should be run by attempting to give a more
    intuitive understanding of how solutes move as a function of their environment.
%    \item The hydrodynamic Ferry model is the simplest pore-flow model, with extensions such
%    as the Donnan-Steric Pore Model which includes extra terms for the pressure gradient
%    across the membrane and membrane charge.
%    \item Dischinger et al. created an empirical correlation to determine the
%    effect of solute physicochemical properties on solute rejection.
  \end{itemize}

@@

  \noindent Amphiphilic molecules are capable of self-assembling into ordered nanostructures.\\

  Lyotropic liquid crystals are a class of amphiphilic molecules that can be cross-linked
  into mechanically strong and highly selective membranes.
  \begin{itemize}
  	\item H\textsubscript{II} phase lyotropic liquid crystals have densely packed, uniform
	sized pores and have the potential to disrupt conventional membrane separation
	techniques by being selective based not only on size and charge, but on chemical
	functionality as well.
	\item Q\textsubscript{I} phase LLCs consist of a tortuous network of 3D interconnected
	pores. They are easier to make.
  \end{itemize}

  We can only learn so much from experiment. MD can give us mechanistic insights with
  atomistic resolution so that we can intelligently design new membranes for 
  solute-specific separations.\\
  %BJC2: simple models from first paragraph fail to adequately describe transport.

  \noindent In our previous work, we studied the transport of 20 small polar molecules
  in an H\textsubscript{II} phase LLC membrane.
  \begin{itemize}
    \item In general, we observed subdiffusive transport behavior characterized by 
    intermittent hops separated by periods of entrapment.
    \item We identified three mechanisms responsible for the solute trapping behavior:
    entanglement among monomer tails, hydrogen bonding with monomer head groups, and
    association with sodium counter ions.
  \end{itemize}

@@

  \noindent The distribution of dwell times is expected to fit a power law
  distribution proportional to $t^{-1-\alpha}$.~\cite{meroz_toolbox_2015}
  \begin{itemize}
	\item Because we are limited to taking measurements at discrete values
	dictated by the output frequency of our simulation trajectories, we fit the
	empirical dwell times to a discrete power law distribution whose maximum
	likelihood $\alpha$ parameter we calculated by maximizing the following
	likelihood function: 
    \begin{equation}
	\mathcal{L}(\beta) = -n\text{ln}\zeta(\beta, x_{min}) -
	\beta\sum_{i=1}^{n} \text{ln}~x_i 
	\label{eqn:powerlaw_likelihood}
	\end{equation}
	where $\beta = 1 + \alpha$, $x_i$ are collected dwell time data points,
	$n$ the total number of data points, and $\zeta$ is the Hurwitz zeta function
	where $x_{min}$ is the smallest measured value of
	$x_i$.~\cite{clauset_power-law_2009} 
	\item We obtained distributions of the hop length standard deviations, $\sigma$, and
	$\alpha$ using bootstrapping.\cite{efron_introduction_1994} 
  \end{itemize}

  In practical applications, the heavy tail of power law distributions can result in 
  arbitrarily long dwell times that are never observed in MD simulations. 
  \begin{itemize}
    \item We can not observe dwell times longer than the length of a single trajectory.
    \item In fact the longest dwell time observed by solutes in this study is X.
    \item In order to compare our anomalous diffusion model to MD results we need to 
    bound the dwell time distribution.
    \item We do this by adding an exponential cut-off to the power law so the dwell 
    time distribution is of the form:
    \begin{equation}
    t^{-1 - \alpha}e^{-\lambda t}
    \end{equation}
    \item We determine MLEs of $\alpha$ and $\lambda$ by maximizing the likelihood function:~\cite{clauset_power-law_2009}
    \begin{equation}
    \mathcal{L}(\alpha, \lambda) = n*(1 - \alpha)ln\lambda - n*ln\Gamma(1 - \alpha, x_{min}\lambda) - \alpha\sum_{i=1}^{n}ln x_i - \lambda\sum_{i=1}^n x_i
    \end{equation}
  \end{itemize}

@@

  \subsection{Markov State Models}\label{method:MSMs}  
  % BJC2: still need a better name
  % name candidates - correlated MSMs, Dynamical mode switching

@@

  In this work, we define a total of 8 discrete states based on the 3 trapping
  mechanisms observed in our previous work. 
  \begin{itemize}
  	\item Therefore, there is no need to subject our trajectories to any reduction
  	in dimensionality. 
  	\item However, we will adopt the validation techniques used in the packages
  	mentioned above.  
  	%BJC: not sure the best way to lay these out
  	\item The states we've chosen include all combinations of trapping mechanisms 
  	in the pore and out of the pore:
        %MRS: ``1. in tails and 5. in pores seem to encompass the things below them. Maybe us ``in tails, not trapped,'' etc. as the last entry in each column.'' 
        %BJC2: good point
    \begin{table}[!htb]
	  \centering
	  \begin{tabular}{ll}
	  \hline
	  1. In tails, not trapped                     & 5. In pores, not trapped                     \\
	  2. In tails and hydrogen bonding             & 6. In pores and hydrogen bonding             \\
	  3. In tails and associated with sodium       & 7. In pores and associated with sodium       \\
	  4. In tails, hydrogen bonding and associated & 8. In pores, hydrogen bonding and associated \\
	  \hline
	  \end{tabular}
	  \caption{}\label{table:water_content}
   \end{table}
%  	\begin{enumerate}
%  	  \item Trapped in tails
%  	  \item Trapped in tails and hydrogen bonding
%  	  \item Trapped in tails and associated with sodium
%  	  \item Trapped in tails, hydrogen bonding and associated
%  	  \item In pores
%  	  \item In pores and hydrogen bonding
%  	  \item In pores and associated with sodium
%  	  \item In pores, hydrogen bonding and associated
%  	\end{enumerate}
  	\item These choices assume that there are no significant kinetic effects resulting
  	from solute conformational changes or pore size fluctuations, an assumption 
  	that may be relaxed in future work.
  \end{itemize}

@@

  % BJC: reduce this figure to solutes studied and put 3 bars for each solute: MD MSD, 
  % one-mode sfbm MSD, two-mode sfbm MSD.
  % BJC2: Actually I think I prefer showing the MSD curves on one plot, like in my research journal. Since it shows shape.
  \begin{figure}
  \centering
  \includegraphics[width=\textwidth]{all_emsds.pdf}
  \caption{}\label{fig:all_msds}
  \end{figure}

@@

  Molecular simulations were performed using the Extreme Science and
  Engineering Discovery Environment (XSEDE), which is supported by National
  Science Foundation grant number ACI-1548562. Specifically, it used the Bridges
  system, which is supported by NSF award number ACI-1445606, at the Pittsburgh
  Supercomputing Center (PSC). This work also utilized the RMACC Summit supercomputer,
  which is supported by the National Science Foundation (awards ACI-1532235 and
  ACI-1532236), the University of Colorado Boulder, and Colorado State
  University. The Summit supercomputer is a joint effort of the University of
  Colorado Boulder and Colorado State University.

>>>>

  % MRS1: I would start with something more central to the content of the paper - how do solutes move in complex separations membranes?  Let people know right from the start what the paper is about.  This paragraph should still be in the intro, but start with something that captures the main question of the paper. 
  % BJC2: This paragraph will need more tweaking but is the gist of what I'm thinking
  % BJC2: I'm not sure if this is exactly what you meant though
%  Making sense of solute motion in complex separations membranes is central to 
%  understanding transport mechanisms and formulating membrane design principles.
  Mathematical descriptions of transport in complex separations membranes are a powerful 
  way to understand mechanisms and formulate membrane design principles.
  % MRS3: seems pretty good.  We can tweak later when turning into text; I don't think that it will really affect the logic of the paper. The introduction will be important here to set the stage, but probably more of that will be in the next paragraph.
  \begin{itemize}
    \item The complexity of a model generally parallels the complexity of the 
    transport mechanism at hand, as well as the desired outputs.
    \item In dense homogeneous membranes, the solution-diffusion model can extract
    diffusion and partition coefficients and has successfully predicted solute transport
    rates.
    \item Analogously, pore-flow models yield predictions of diffusion coefficients
    and solute transport rates in nanoporous membranes.~\cite{wijmans_solution-diffusion_1995}
    \item Single particle tracking allows us to further characterize more complex
    diffusive behavior.
    \item Finally, molecular simulation provides a framework for studying both single
    particle dynamics and bulk transport properties with atomic resolution.
    \item All of these approaches facilitate speculation on the molecular origins of
    separations and suggest experiments that should be run by attempting to give a more
    intuitive understanding of how solutes move as a function of their environment.
%    \item The hydrodynamic Ferry model is the simplest pore-flow model, with extensions such
%    as the Donnan-Steric Pore Model which includes extra terms for the pressure gradient
%    across the membrane and membrane charge.
%    \item Dischinger et al. created an empirical correlation to determine the
%    effect of solute physicochemical properties on solute rejection.
  \end{itemize}

@@

  %MRS3: I'd omit next line (That I commented out); you talk about membranes above, so better to start the next paragraph talking about membranes.
  %\noindent Amphiphilic molecules are capable of self-assembling into ordered nanostructures.\\

  Lyotropic liquid crystals are a class of amphiphilic molecules that can be cross-linked
  into mechanically strong and highly selective membranes.
  \begin{itemize}
  	\item H\textsubscript{II} phase lyotropic liquid crystals have densely packed, uniform
	sized pores and have the potential to disrupt conventional membrane separation
	techniques by being selective based not only on size and charge, but on chemical
	functionality as well.
	\item Q\textsubscript{I} phase LLCs consist of a tortuous network of 3D interconnected
	pores. They are easier to make.
        %MRS3: probably want to put a more here about why we are studing these membranes, and what advantages they might have, and why transport in them is complicated.  
  \end{itemize}

  %MRS3: One missing connection between the above information and what is in the paper would be making the connection between MD simulations
  %and mathematical modeling, using MD to provide/justify the fuctional form and/or some of the parameters for the model, 
  %rather than just experiment.

  %MRS3: a suggested tweak in language
  %We can only learn so much from experiment. 
  There are limitations to what we can learn from experiment about the atomistic details of transport.
  MD can give us mechanistic insights with
  atomistic resolution so that we can intelligently design new membranes for 
  solute-specific separations.\\
  %BJC2: simple models from first paragraph fail to adequately describe transport.

  \noindent In our previous work, we studied the transport of 20 small polar molecules
  in an H\textsubscript{II} phase LLC membrane.
  \begin{itemize}
    \item In general, we observed subdiffusive transport behavior characterized by 
    intermittent hops separated by periods of entrapment.
    \item We identified three mechanisms responsible for the solute trapping behavior:
    entanglement among monomer tails, hydrogen bonding with monomer head groups, and
    association with sodium counter ions.
    %MRS3: point out that they were all subdiffusive (unlike water)
  \end{itemize}

@@

  \noindent The distribution of dwell times is expected to fit a power law
  distribution proportional to $t^{-1-\alpha}$.~\cite{meroz_toolbox_2015}
  \begin{itemize}
	\item Because we are limited to taking measurements at discrete values
	dictated by the output frequency of our simulation trajectories, we fit the
	empirical dwell times to a discrete power law distribution whose maximum
	likelihood $\alpha$ parameter we calculated by maximizing the following
	likelihood function: 
    \begin{equation}
      %MRS: use \ln instead of ln or \text{ln} 
	\mathcal{L}(\beta) = -n\ln \zeta(\beta, x_{min}) -
	\beta\sum_{i=1}^{n} \ln x_i 
	\label{eqn:powerlaw_likelihood}
	\end{equation}
	where $\beta = 1 + \alpha$, $x_i$ are collected dwell time data points,
	$n$ the total number of data points, and $\zeta$ is the Hurwitz zeta function
	where $x_{min}$ is the smallest measured value of
	$x_i$.~\cite{clauset_power-law_2009} 
	\item We obtained distributions of the hop length standard deviations, $\sigma$, and
	$\alpha$ using bootstrapping.\cite{efron_introduction_1994} 
  \end{itemize}

  In practical applications, the heavy tail of power law distributions can result in 
  arbitrarily long dwell times that are never observed in MD simulations. 
  \begin{itemize}
    \item We can not observe dwell times longer than the length of a single trajectory.
    \item In fact the longest dwell time observed by solutes in this study is X.
    \item In order to compare our anomalous diffusion model to MD results we need to 
    bound the dwell time distribution.
    \item We do this by adding an exponential cut-off to the power law so the dwell 
    time distribution is of the form:
    \begin{equation}
    t^{-1 - \alpha}e^{-\lambda t}
    \end{equation}
    \item We determine MLEs of $\alpha$ and $\lambda$ by maximizing the likelihood function:~\cite{clauset_power-law_2009}
    \begin{equation}
    %MRS3: I replaced ln with \ln
    \mathcal{L}(\alpha, \lambda) = n(1 - \alpha)\ln\lambda - n\ln\Gamma(1 - \alpha, x_{min}\lambda) - \alpha\sum_{i=1}^{n}\ln x_i - \lambda\sum_{i=1}^n x_i
    \end{equation}
  \end{itemize}

@@

  \subsection{Markov State Models}\label{method:MSMs}  
  % BJC2: still need a better name
  % name candidates - correlated MSMs, Dynamical mode switching
  % MRS3: probably not ``correlated MSMs'', that's insufficiently descriptive. Many things dynamically change modes.
  % Maybe ``Markov state-dependent dynamics?'' You have dynamics/transport that are a function of the state, which is 
  % determined by a Markov state model.

@@

  In this work, we define a total of 8 discrete states based on the 3 trapping
  mechanisms observed in our previous work. 
  \begin{itemize}
  	\item Therefore, there is no need to subject our trajectories to any reduction
  	in dimensionality. 
  	\item However, we will adopt the validation techniques used in the packages
  	mentioned above.  
  	%BJC: not sure the best way to lay these out
  	\item The states we've chosen include all combinations of trapping mechanisms 
  	in the pore and out of the pore:
        %MRS: ``1. in tails and 5. in pores seem to encompass the things below them. Maybe us ``in tails, not trapped,'' etc. as the last entry in each column.'' 
        %BJC2: good point
    \begin{table}[!htb]
	  \centering
          %MRS3: tweaked formatting, change back if you want!
	  \begin{tabular}{|c|c|}
	  \hline
	  1. In tails, not trapped                     & 5. In pores, not trapped                     \\
	  2. In tails and hydrogen bonding             & 6. In pores and hydrogen bonding             \\
	  3. In tails and associated with sodium       & 7. In pores and associated with sodium       \\
	  4. In tails, hydrogen bonding and associated & 8. In pores, hydrogen bonding and associated \\
	  \hline
	  \end{tabular}
	  \caption{}\label{table:water_content}
   \end{table}
%  	\begin{enumerate}
%  	  \item Trapped in tails
%  	  \item Trapped in tails and hydrogen bonding
%  	  \item Trapped in tails and associated with sodium
%  	  \item Trapped in tails, hydrogen bonding and associated
%  	  \item In pores
%  	  \item In pores and hydrogen bonding
%  	  \item In pores and associated with sodium
%  	  \item In pores, hydrogen bonding and associated
%  	\end{enumerate}
  	\item These choices assume that there are no significant kinetic effects resulting
  	from solute conformational changes or pore size fluctuations, an assumption 
  	that may be relaxed in future work.
  \end{itemize}

@@

  % BJC: reduce this figure to solutes studied and put 3 bars for each solute: MD MSD, 
  % one-mode sfbm MSD, two-mode sfbm MSD.
  % BJC2: Actually I think I prefer showing the MSD curves on one plot, like in my research journal. Since it shows shape.
  % MRS3: OK.
  \begin{figure}
  \centering
  \includegraphics[width=\textwidth]{all_emsds.pdf}
  \caption{}\label{fig:all_msds}
  \end{figure}

@@

  %MRS3: add specific info for PRF and GAANN
  This work was supported in part by the ACS Petroleum Research Fund grant XXXXXX and GAANN grant YYYYY.
  Molecular simulations were performed using the Extreme Science and
  Engineering Discovery Environment (XSEDE), which is supported by National
  Science Foundation grant number ACI-1548562. Specifically, it used the Bridges
  system, which is supported by NSF award number ACI-1445606, at the Pittsburgh
  Supercomputing Center (PSC). This work also utilized the RMACC Summit supercomputer,
  which is supported by the National Science Foundation (awards ACI-1532235 and
  ACI-1532236), the University of Colorado Boulder, and Colorado State
  University. The Summit supercomputer is a joint effort of the University of
  Colorado Boulder and Colorado State University.
